---
title: "Github copy"
output: html_notebook
---


Load Libraries
```{r}
library(lavaan)
library(haven)
library(psych)
library(GPArotation)
library(moments)
```
Load Data
Clean Data, make sure your data has columns representing each question
```{r}
data2<-read_dta("C:/Users/Downloads/data2.dta")
```
CFA on Self Determination
Already have a theory on grouping so we are testing it
```{r}
selfDeterminationModel <- '
  autonomyLatent =~ Aut_1+Aut_2+Aut_3+Aut_4
  competenceLatent =~ Com_1+Com_2+Com_3+Com_4
  relatednessLatent =~ Rel_1+Rel_2+Rel_3+Rel_4
'

fit <- cfa(selfDeterminationModel, data=data2)
summary(fit, fit.measures=TRUE)
```
Cronbach Alpha on Self Determination
```{r}
# Cronbach's Alpha for Each Subscale
alpha_autonomy <- alpha(data2[, c("Aut_1", "Aut_2", "Aut_3", "Aut_4")]) 
alpha_autonomy 
alpha_competence <- alpha(data2[, c("Com_1", "Com_2", "Com_3", "Com_4")])
alpha_competence
alpha_relatedness <- alpha(data2[, c("Rel_1", "Rel_2", "Rel_3", "Rel_4")])
alpha_relatedness
alpha_basic_needs <- alpha(data2[, c("Aut_1", "Aut_2", "Aut_3", "Aut_4", "Com_1", "Com_2", "Com_3", "Com_4", "Rel_1", "Rel_2", "Rel_3", "Rel_4")])
alpha_basic_needs
```
EFA for Emotional Experiences
Two theories for groupings, 3 groups (Enjoyment, boredom, anxiety) or 2 groups (enjoyment and negative emotion)
USed EFA, scree plot, and parallel analysis to test if 2 or 3 factors best represents the data.
Two of anxiety items did not lead well so they were removed
```{r}
data_for_efa <- data2[, c("Enj_1","Enj_2","Enj_3","Enj_4","Bor_1","Bor_2","Bor_3","Anx_1","Anx_2","Anx_3")]
data_for_efa <- na.omit(data_for_efa)
scree(data_for_efa)
fa.parallel(data_for_efa, fa = "fa")
efa_result <- factanal(data_for_efa, factors = 2, rotation = "varimax")
print(efa_result)

EmotionModel <- '
  enjoymentLatent =~ Enj_1+Enj_2+Enj_3+Enj_4
  boredomLatent =~ Bor_1+Bor_2+Bor_3
  anxietyLatent =~ Anx_1+Anx_2+Anx_3
'

fit <- cfa(EmotionModel, data=data2)
summary(fit, fit.measures=TRUE)
```
CFA on Emotional Expereinces to test the 2 factor model
```{r}
cor(data2[c("Enj_1","Enj_2","Enj_3","Enj_4","Bor_1","Bor_2","Bor_3","Anx_1","Anx_2","Anx_3")], use = "complete.obs")
EmotionModel <- '
  enjoymentLatent =~ Enj_1+Enj_2+Enj_3+Enj_4
  boredomAnxietyLatent =~ Bor_1+Bor_2+Bor_3+Anx_3
'

fit <- cfa(EmotionModel, data=data2)
summary(fit, fit.measures=TRUE)
```
Cronbach ALpha on Emotional Experiences
```{r}
# Calculating Cronbach's Alpha for Emotional Experiences
alpha_enjoyment <- alpha(data2[, c("Enj_1", "Enj_2", "Enj_3", "Enj_4")])
alpha_enjoyment 
alpha_anxiety <- alpha(data2[, c("Anx_1", "Anx_2", "Anx_3")])
alpha_anxiety 
alpha_boredom <- alpha(data2[, c("Bor_1", "Bor_2", "Bor_3")])
alpha_boredom 
alpha_emotional <- alpha(data2[, c("Enj_1", "Enj_2", "Enj_3", "Enj_4", "Anx_1", "Anx_2", "Anx_3", "Bor_1", "Bor_2", "Bor_3")],check.keys=TRUE)
alpha_emotional

# Regrouping
alpha_anxietyBoredom <- alpha(data2[, c( "Anx_3", "Bor_1", "Bor_2", "Bor_3")])
alpha_anxietyBoredom 
alpha_emotionalRemovals <- alpha(data2[, c("Enj_1", "Enj_2", "Enj_3", "Enj_4", "Anx_3", "Bor_1", "Bor_2", "Bor_3")],check.keys=TRUE)
alpha_emotionalRemovals 

all_responses <- c(data2$Anx_3, data2$Bor_1, data2$Bor_2, data2$Bor_3)
sd(all_responses, na.rm = T)
skewness(all_responses, na.rm=T)
kurtosis(all_responses, na.rm = T)
```
CFA on Motivation Model
testing existing theory
```{r}
cor(data2[c("Int_1"+"Int_2","Int_3"+"Int_4","Self_1"+"Self_2"+"Self_3"+"Self_4""Ext_1"+"Ext_2"+"Ext_3"+"Ext_4""Amo_1"+"Amo_2"+"Amo_3"+"Amo_4")], use = "complete.obs")
MotivationModel <- '
  intrinsicLatent =~ Int_1+Int_2+Int_3+Int_4
  introjectLatent =~ Self_1+Self_2+Self_3+Self_4
  externalLatent =~ Ext_1+Ext_2+Ext_3
  amotivatedLatent =~ Amo_1+Amo_2+Amo_3+Amo_4
'
fit <- cfa(MotivationModel, data=data2)
summary(fit, fit.measures=TRUE)
```
EFA on motivation (given that intrinsic seemed to load with identified)
tried 3 and 4 factor model, went with 3
```{r}
data_for_efa <- data2[, c("Int_1","Int_2","Int_3","Int_4","Self_1","Self_2","Self_3","Self_4","Ext_1","Ext_2","Ext_3","Ext_4","Amo_1","Amo_2","Amo_3","Amo_4")]
data_for_efa <- na.omit(data_for_efa)
scree(data_for_efa)
fa.parallel(data_for_efa, fa = "fa")
efa_result <- factanal(data_for_efa, factors = 4, rotation = "varimax")
print(efa_result)
alpha(data2[c("Int_1","Int_2","Int_3","Int_4")])

MotivationModel <- '
  intrinsicLatent =~ Int_1+Int_2+Int_3+Int_4+Self_1+Self_2+Self_3+Self_4
  externalLatent =~ Ext_1+Ext_2+Ext_3
  amotivatedLatent =~ Amo_1+Amo_2+Amo_3+Amo_4
'
fit <- cfa(MotivationModel, data=data2)
summary(fit, fit.measures=TRUE)
```
Cronbach alpha Motivation
```{r}
# Calculating Cronbach's Alpha for Motivation
alpha_intrinsic <- alpha(data2[, c("Int_1", "Int_2", "Int_3", "Int_4")])
alpha_intrinsic 
alpha_self <- alpha(data2[, c("Self_1", "Self_2", "Self_3", "Self_4")])
alpha_self 
alpha_positiveEmotion <- alpha(data2[, c("Int_1", "Int_2", "Int_3", "Int_4", "Self_1", "Self_2", "Self_3", "Self_4")])
alpha_positiveEmotion 
alpha_external <- alpha(data2[, c("Ext_1", "Ext_2", "Ext_3", "Ext_4")])
alpha_external 
alpha_externalRemoval <- alpha(data2[, c("Ext_1", "Ext_2", "Ext_3")])
alpha_externalRemoval 
alpha_amotivated <- alpha(data2[, c("Amo_1", "Amo_2", "Amo_3", "Amo_4")])
alpha_amotivated 
alpha_motivation <- alpha(data2[, c("Int_1", "Int_2", "Int_3", "Int_4", "Self_1", "Self_2", "Self_3", "Self_4", "Ext_1", "Ext_2", "Ext_3", "Ext_4", "Amo_1", "Amo_2", "Amo_3", "Amo_4")], check.keys=TRUE)
alpha_motivation 
alpha_motivationRemoval <- alpha(data2[, c("Int_1", "Int_2", "Int_3", "Int_4", "Self_1", "Self_2", "Self_3", "Self_4", "Ext_1", "Ext_2", "Ext_3", "Amo_1", "Amo_2", "Amo_3", "Amo_4")], check.keys=TRUE)
alpha_motivationRemoval 

all_responses <- c(data2$Int_1, data2$Int_2, data2$Int_3, data2$Int_4, data2$Self_1, data2$Self_2, data2$Self_3, data2$Self_4)
all_responses <- c(data2$Ext_1, data2$Ext_2, data2$Ext_3)

mean(all_responses, na.rm = T)
sd(all_responses, na.rm = T)
skewness(all_responses, na.rm=T)
kurtosis(all_responses, na.rm = T)
```
Implementation Model
EFA to explore grouping, ran 3 and 4 factors, 3 fit best
```{r}
data_for_efa <- data2[, c("Imp_1","Imp_2","Imp_3","Imp_4","Imp_5","Imp_6","Imp_7","Imp_8","Imp_9","Imp_10","Imp_11","Imp_12")]
data_for_efa <- na.omit(data_for_efa)
scree(data_for_efa)
fa.parallel(data_for_efa, fa = "fa")
efa_result <- factanal(data_for_efa, factors = 3, rotation = "varimax")
print(efa_result)
```
Cronbach alpha Implementation
```{r}
# Cronbach's Alpha for Implementation
alpha_implementation <- alpha(data2[, c("Imp_1", "Imp_2", "Imp_3", "Imp_4", "Imp_5", "Imp_6", "Imp_7", "Imp_8", "Imp_9", "Imp_10", "Imp_11", "Imp_12")])
alpha_implementation 

# Identify
alpha_identify <- alpha(data2[, c("Imp_1", "Imp_2", "Imp_3", "Imp_4")])
alpha_identify
# Inquire/Assess
alpha_inquire <- alpha(data2[, c("Imp_5", "Imp_6", "Imp_7", "Imp_8")])
alpha_inquire
# Manage
alpha_manage <- alpha(data2[, c("Imp_9", "Imp_10", "Imp_11", "Imp_12")])
alpha_manage 
```
CFA Implementation
```{r}
SBTAModel <- '
  identifyLatent =~ Imp_1+Imp_2+Imp_3+Imp_4
  inquireLatent =~ Imp_5+Imp_6+Imp_7+Imp_8
  manageLatent =~ Imp_9+Imp_10+Imp_11+Imp_12
'
fit <- cfa(SBTAModel, data=data2)
summary(fit, fit.measures=TRUE)
```
SEM Model - Full model
```{r}
# SEM model specification
semModel <- '
  # Measurement Model
  autonomyLatent =~ Aut_1 + Aut_2 + Aut_3 + Aut_4
  competenceLatent =~ Com_1 + Com_2 + Com_3 + Com_4
  relatednessLatent =~ Rel_1 + Rel_2 + Rel_3 + Rel_4
  
  enjoymentLatent =~ Enj_1 + Enj_2 + Enj_3 + Enj_4
  boredomAnxietyLatent =~ Bor_1 + Bor_2 + Bor_3 + Anx_3

  intrinsicLatent =~ Int_1+Int_2+Int_3+Int_4+Self_1+Self_2+Self_3+Self_4
  externalLatent =~ Ext_1 + Ext_2 + Ext_3
  amotivatedLatent =~ Amo_1 + Amo_2 + Amo_3 + Amo_4

  identifyLatent =~ Imp_1 + Imp_2 + Imp_3 + Imp_4
  inquireLatent =~ Imp_5 + Imp_6 + Imp_7 + Imp_8
  manageLatent =~ Imp_9 + Imp_10 + Imp_11 + Imp_12

  # Structural Model
  intrinsicLatent ~ autonomyLatent + competenceLatent + relatednessLatent + enjoymentLatent + boredomAnxietyLatent
  externalLatent ~ autonomyLatent + competenceLatent + relatednessLatent + enjoymentLatent + boredomAnxietyLatent
  amotivatedLatent ~ autonomyLatent + competenceLatent + relatednessLatent + enjoymentLatent + boredomAnxietyLatent
  
  identifyLatent ~ intrinsicLatent + externalLatent + amotivatedLatent 
  inquireLatent ~ intrinsicLatent + externalLatent + amotivatedLatent 
  manageLatent ~ intrinsicLatent + externalLatent + amotivatedLatent 

  # Covariance Specifications
  
  intrinsicLatent ~~ amotivatedLatent
  intrinsicLatent ~~ externalLatent
  externalLatent ~~ amotivatedLatent
  
  autonomyLatent ~~ competenceLatent
  autonomyLatent ~~ relatednessLatent
  competenceLatent ~~ relatednessLatent

  enjoymentLatent ~~ boredomAnxietyLatent

  identifyLatent ~~ inquireLatent
  identifyLatent ~~ manageLatent
  inquireLatent ~~ manageLatent

'

# Fit the model
fit <- sem(semModel, data=data2)

# Summary of the fit
summary(fit, fit.measures=TRUE)
```
Separate Roles
Create RoleComparison variable where Leadership = 1 and Practitioners=2
```{r}
# Step 1: Create the RoleComparison column
data2 <- data2 %>%
  mutate(RoleComparison = case_when(
    Role %in% c(1,2,6,7) ~ 1,
    Role %in% c(8,9,5,10,11,12) ~ 2,
    TRUE ~ NA_real_ # Assign NA for rows that don't meet the conditions
  ))

# Step 2: Filter out rows with NA in RoleComparison and rename the dataset
dataRole <- data2 %>%
  filter(!is.na(RoleComparison))

table(dataRole$RoleComparison)
```
Create 2 dataframes,one with leadership another with student support
```{r}
# Create dataframe where RoleComparison is 1
dataRole1 <- dataRole %>%
  filter(RoleComparison == 1)

# Create dataframe where RoleComparison is 2
dataRole2 <- dataRole %>%
  filter(RoleComparison == 2)
```
Test the Two Roles
```{r}
# Fit the model
fit1 <- sem(semModel, data=dataRole1)

# Summary of the fit
summary(fit1, fit.measures=TRUE)
# Fit the model
fit2 <- sem(semModel, data=dataRole2)

# Summary of the fit
summary(fit2, fit.measures=TRUE)
```

